DESING AND SIMULATION SOFTWARE FOR RHODOTRON TYPE ELECTRON ACCELERATORS

### 2023-08-25 subsections/design/cavity/cavity.tex
25/08/2023 update

The first and the most important design parameter for a cavity is the operating RF frequency.
After an operating RF frequency is set and the desired $R1/R2$ relation \fromfig{pottier_table1} is selected, design parameters of acceleration plane of the cavity is fully determined.

>>>>

The first and the most important design parameter for a cavity is the operating RF frequency.
After an operating RF frequency is set and the desired $R1/R2$ relation in \fromfig{pottier_table1} is selected, design parameters of acceleration plane of the cavity is fully determined.

### 2023-08-25 subsections/future/future_work.tex
25/08/2023 update

\textit{Rhodotron Simulation}, which has been the main focus of this thesis, is ready and waiting to be tested with the cavity that is being manufactured; 
several improvements and new feature implementations are underway in the mean time. 
They include the following;
\begin{figure}[H]
    \begin{subfigure}[t]{0.5\textwidth}
        \begin{itemize}
            \item Electric and magnetic field import
            \item Field generator module to directly produce field files from specified cavity
            \item Synchrotron radiation calculations to determine radiation and energy loss
            \item \ee interactions
            \item Redesign of the GUI
            \item 3D render in GUI
        \end{itemize}
    \end{subfigure}
    \begin{subfigure}[t]{0.5\textwidth}
        \begin{itemize}
            \item Refactoring of the \eEM interaction and logging to improve performance
            \item Refactoring of the GUI Render Frame to improve render speed
            \item Refactoring the magnet class to introduce field leaks
            \item Extention of analysis tools in Analyze Frame
            \item $L_{out}$ sweep in Sweep Frame
        \end{itemize}
    \end{subfigure}
\end{figure}
After the cavity manufacturing is concluded, energy gain predictions of this software will be tested on this cavity. 
Complementary bending magnet design will be the first challange of \textit{Rhodotron Simulation}, which will provide tremendous testing and improvement opportunuties.

>>>>

\textit{Rhodotron Simulation}, which has been the main focus of this thesis, is ready and waiting to be tested with the cavity that is being manufactured; 
several improvements and new feature implementations are underway in the mean time. 
They include the following;
\begin{figure}[H]
    \begin{subfigure}[t]{0.5\textwidth}
        \begin{itemize}
            \item Electric and magnetic field import
            \item Field generator module to directly produce field files from specified cavity
            \item Synchrotron radiation calculations to determine radiation and energy loss
            \item \ee interactions
            \item Redesign of the GUI
            \item 3D render in GUI
        \end{itemize}
    \end{subfigure}
    \begin{subfigure}[t]{0.5\textwidth}
        \begin{itemize}
            \item Refactoring of the \eEM interaction and logging to improve performance
            \item Refactoring of the GUI Render Frame to improve render speed
            \item Refactoring the magnet class to introduce field leaks
            \item Extention of analysis tools in Analyze Frame
            \item $L_{out}$ sweep in Sweep Frame
        \end{itemize}
    \end{subfigure}
\end{figure}
After the cavity manufacturing is concluded, energy gain predictions of this software will be tested on this cavity. 
Complementary bending magnet design will be the first challange of \textit{Rhodotron Simulation}, which will provide tremendous testing and improvement opportunities.

### 2023-08-25 subsections/simulation/gui/gui.tex
25/08/2023 update

\subsection{Simulation Frame}
This frame spawns and manages the \textit{simulation engine}, configures and starts the simulation. 
It has a progress bar that shows the current progress of the simulation, communucating with \textit{UI handler thread} in \textit{simulation engine}.
Simulation frame after a simulation run has been completed can be observed in \fromfig{gui_simulate_1} of \fromapp{supp}.

\subsection{Render Frame}
Since the rendering capabilities of \textit{ROOT} is superior than \textit{gnuplot}, the user can render a playback of the simulation in real time, 
see a specific time frame, export as snapshot or animated gif file.
Two example renders can be seen in \fromfig{gui_render_1} and \fromfig{gui_render_2} of \fromapp{supp}.
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{../../../figures/rhodoSim/GUI_render_frame_1.png}
    \caption{Render frame of \textit{GUI}.}
    \label{fig:gui_render_1}
\end{figure}

\subsection{Analyze Frame}
Analyze frame provides tools for analyzing and visualizing the simulation data. 
In the current version, \textbf{E} distribution histogram and \textbf{E(t)} graph of each electron are implementated into this frame.
This frame is a work in progress and will be the focus of improvement and become a really powerfull tool in the near future.
In \fromfig{gui_analyze_Edist}, \textit{Analyze Frame} plotting \textbf{E} distribution histogram, and in \fromfig{gui_analyze_Et} and \fromfig{gui_analyze_Et2} of \fromapp{supp}, \textit{Analyze Frame} plotting \textbf{E(t)} graphs can be observed.
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{../../../figures/rhodoSim/GUI_analyze_Edist.png}
    \caption{Analyze frame of \textit{GUI} E distribution.}
    \label{fig:gui_analyze_Edist}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{../../../figures/rhodoSim/GUI_analyze_Et_1.png}
    \caption{Analyze frame of \textit{GUI} E(t).}
    \label{fig:gui_analyze_Et}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{../../../figures/rhodoSim/GUI_sweep_muE.png}
    \caption{Sweep frame of \textit{GUI} \\ $\phi_{lag}$ $\mu E$ result.}
    \label{fig:gui_sweep_muE}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{../../../figures/rhodoSim/GUI_sweep_sE.png}
    \caption{Sweep frame of \textit{GUI} \\ $\phi_{lag}$ sweep $\sigma E$ result.}
    \label{fig:gui_sweep_sE}
\end{figure}

>>>>

\subsection{Simulation Frame}
This frame spawns and manages the \textit{simulation engine}, configures and starts the simulation. 
It has a progress bar that shows the current progress of the simulation, communucating with \textit{UI handler thread} in \textit{simulation engine}.
Simulation frame while a simulation is running can be observed in \fromfig{gui_simulate_1} of \fromapp{supp}.

\subsection{Render Frame}
Since the rendering capabilities of \textit{ROOT} is superior than \textit{gnuplot}, the user can render a playback of the simulation in real time, 
see a specific time frame, export as snapshot or animated gif file.
Two example renders can be seen in \fromfig{gui_render_1} and \fromfig{gui_render_2} of \fromapp{supp}.
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{../../../figures/rhodoSim/GUI_render_frame_5.png}
    \caption{Render frame of \textit{GUI}.}
    \label{fig:gui_render_1}
\end{figure}

\subsection{Analyze Frame}
Analyze frame provides tools for analyzing and visualizing the simulation data. 
In the current version, \textbf{E} distribution histogram and \textbf{E(t)} graph of each electron are implementated into this frame.
This frame is a work in progress and will be the focus of improvement and become a really powerfull tool in the near future.
In \fromfig{gui_analyze_Edist}, \textit{Analyze Frame} plotting \textbf{E} distribution histogram, and in \fromfig{gui_analyze_Et} and \fromfig{gui_analyze_Et2} of \fromapp{supp}, \textit{Analyze Frame} plotting \textbf{E(t)} graphs can be observed.
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{../../../figures/rhodoSim/GUI_analyze_Edist_2.png}
    \caption{Analyze frame of \textit{GUI} E distribution.}
    \label{fig:gui_analyze_Edist}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{../../../figures/rhodoSim/GUI_analyze_Et_3.png}
    \caption{Analyze frame of \textit{GUI} E(t).}
    \label{fig:gui_analyze_Et}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{../../../figures/rhodoSim/GUI_sweep_muE_3.png}
    \caption{Sweep frame of \textit{GUI} \\ $\phi_{lag}$ $\mu E$ result.}
    \label{fig:gui_sweep_muE}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{../../../figures/rhodoSim/GUI_sweep_sE_3.png}
    \caption{Sweep frame of \textit{GUI} \\ $\phi_{lag}$ sweep $\sigma E$ result.}
    \label{fig:gui_sweep_sE}
\end{figure}

### 2023-08-25 subsections/simulation/intermediate_versions/intermediate.tex
25/08/2023 update

    \begin{subfigure}{0.8\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/analiz/mag_lf_rk2_dt-Tsim.png}
        \caption*{$dt$ vs $T_{sim}$}
    \end{subfigure}
    \caption{Comparing Leap-frog, RK4-2 performance on \eB interaction\\ $E_{in}=1$MeV, \textbf{B}=$0.1$T, $t_{end}=5$ns}
    \label{fig:mag_lf_rk2_comparison}
\end{figure}
The data from these tests can be found in \fromtabs{lf_mag_table}{rk2_mag_table} of \fromapp{data_graph}.
To investigate the data further, one can define a performance measurement, $F$.
\begin{eqnarray*}
    F &\propto& 1/T \\
    F &\propto& 1/\Delta E 
\end{eqnarray*}
When $dt=10^-5$ns was taken as reference point due to providing a good balance of accuracy and computational intensity,
\begin{eqnarray} \label{eq:f_lf_rk_mag_1}
    \Delta E_{LF} &=& 110 \times 10^{-5} MeV \nonumber\\
    \Delta E_{RK}^1 &=& 64 \times 10^{-5} MeV \nonumber\\
    T_{LF}^1 &=& 9.44 \pm 0.03 s \nonumber\\
    T_{RK}^1 &=& 21.33 \pm 0.02 s \nonumber\\
    \Delta E_{LF} \times T_{LF} &=& 104 \times 10^{-4} \pm 10^{-4} MeV s \nonumber\\ %0.003
    \Delta E_{RK} \times T_{RK} &=& 137 \times 10^{-4} \pm 2\times 10^{-4}  MeV s \nonumber\\ %0.001
    F_{LF}/F_{RK} &=& \frac{\Delta E_{RK} \times T_{RK} }{\Delta E_{LF} \times T_{LF}^1} = 1.32 \pm 0.01
\end{eqnarray}
Also, observing from the data,
\begin{eqnarray} \label{eq:f_lf_rk_mag_2}
    \Delta E_{LF} (dt=3 \times 10^{-5}) \approx \Delta E_{RK} (dt=5 \times 10^{-5}) &\approx& 32.5 \times 10^{-4} MeV \nonumber\\
    T_{LF}(dt=3 \times 10^{-5})  &=& 3.18 \pm 0.02 s \nonumber\\ %0.0032
    T_{RK}(dt=5 \times 10^{-5})  &=& 4.30 \pm 0.01 s \nonumber\\ %0.0022
    F_{LF}/F_{RK} = \frac{T_{RK}(dt=5 \times 10^{-5})}{T_{LF}(dt=3 \times 10^{-5})} &=& 1.4 \pm 0.1
\end{eqnarray}
Uncertainty of \fromeq{f_lf_rk_mag_2} was taken high due to the approximation. After combining \fromeqs{f_lf_rk_mag_1}{f_lf_rk_mag_2}, $F$ can be calculated as
\begin{equation}
    F_{LF}/F_{RK} = 1.36 \pm 0.05
\end{equation}
Therefore, \textit{Leap-frog} was found to be the better choice as it provided with $1.36 \pm 0.05$ times accuracy in \eB interactions at a given time with respect to \textit{RK4}. However,
\textit{RK4} was promising in situations where decreasing the stepsize, $dt$, is not viable. Therefore both were integrated into \textit{Rhodotron Simulation} for the user to decide.
\textit{RK4-2} renders from these test can be found in \fromfig{mag_rk_render} of \fromapp{example_simulation_runs}.

@@

Considering the $\vecthreeBF{E}$ is parallel to the beam path, potential difference $V$ in the trejectory until \vecNum{-0.753}{0}{0}m is
\begin{eqnarray}
    \Delta V^1 &=& - \int \dotProdBF{E}{ds} \\
             &=& - \int_{-0.753}^{0.753} -2.65616 \times dx \\
             &=& 2.65616 \times 1.506 \\
    \Delta V^1 &=& 4 MV \\
    \Delta E^1 &=& 4 MeV \\
    E_{exitTH}^1 &=& 5 MeV
\end{eqnarray}
In the second test on the other hand, the beam would be injected into a different static uniform electric field, \\
$\vecthreeBF{E}=(0, -5.31232, 0)$ MV/m where $-0.753<x<0.753$ and $-0.753<y<0.753$, $\vecthreeBF{E}=0$ elsewhere.
\begin{eqnarray}
    \Delta V^2   &=& - \int \dotProdBF{E}{ds} \\
                 &=& - \int_{0}^{0.753} -5.31232 \times dy \\
                 &=& 5.31232 \times 0.753 \\
    \Delta V^2   &=& 4 MV \\
    \Delta E^2   &=& 4 MeV \\
    E_{exitTH}^2 &=& 5 MeV
\end{eqnarray}
Therefore, in the both tests, the beam was expected to exit $\vecthreeBF{E}$ with $E_{exitTH} = 5$ MeV. 
To also measure the variance in simulation completion times, $T_{sim}$, set of $10$ runs were completed at the configuration for each $dt$ value.
Results from these tests can be observed in \fromfig{lf_rk2_perp_stat_E_comparison} and \fromfig{lf_rk2_par_stat_E_comparison} of \fromapp{data_graph}.
\begin{figure}[H]
    \centering
    \begin{subfigure}{0.9\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/statE_par.png}
        \caption*{Setup-1, $\textbf{V}_{\parallel}$=4MV}
    \end{subfigure}
    \begin{subfigure}{0.9\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/statE_perp.png}
        \caption*{Setup-2, $\textbf{V}_{\perp}$=4MV}
    \end{subfigure}
    \caption{Render of the test setups. \\ $E_{in}=1$ MeV, $t_{end}=6$ns}
\end{figure}

@@

    \begin{subfigure}{0.9\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/analiz/90staticE_lf_rk2_dt-Tsim.png}
    \end{subfigure}
    \caption{Comparing Leap-frog, RK4-2 performance on \eE interaction\\ $E_{in}=1$MeV, $\textbf{V}_{\perp}$=4MV, $t_{end}=6$ns}
    \label{fig:lf_rk2_perp_stat_E_comparison}
\end{figure}
Taking the $dt=10^{-5}$ns for the reference point as before, the relative performance can be calculated.
\begin{eqnarray*}
    \begin{aligned}
        \Delta E_{LF}^1 &=& 22 \times 10^{-6} MeV \\
        \Delta E_{RK}^1 &=& 16 \times 10^{-6} MeV \\
        T_{LF}^1 &=& 12.40 \pm 0.04 s \\
        T_{RK}^1 &=& 40.08 \pm 0.12 s \\
    \end{aligned}
    \qquad
    \begin{aligned}
        \Delta E_{LF}^2 &=& 15 \times 10^{-5} MeV \\
        \Delta E_{RK}^2 &=& 9.3 \times 10^{-5} MeV \\
        T_{LF}^2 &=& 10.75 \pm 0.04 s \\
        T_{RK}^2 &=& 34.93 \pm 0.39 s 
    \end{aligned}
\end{eqnarray*}

>>>>

    \begin{subfigure}{0.8\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/analiz/mag_lf_rk2_dt-Tsim.png}
        \caption*{$dt$ vs $T_{sim}$}
    \end{subfigure}
    \caption{Comparing Leap-frog, RK4-2 performance on \eB interaction\\ $E_{in}=1$MeV, \textbf{B}=$0.1$T, $t_{end}=5$ns}
    \label{fig:mag_lf_rk2_comparison}
\end{figure}
The data from these tests can be found in \fromtabs{lf_mag_table}{rk2_mag_table} of \fromapp{data_graph}.
To investigate the data further, one can define a performance measurement, $F$.
\begin{eqnarray*}
    F &\propto& 1/T \\
    F &\propto& 1/\Delta E 
\end{eqnarray*}
When $dt=10^{-5}$ ns was taken as reference point due to providing a good balance of accuracy and computational intensity,
\begin{eqnarray} \label{eq:f_lf_rk_mag_1}
    \Delta E_{LF} &=& 110 \times 10^{-5} MeV \nonumber\\
    \Delta E_{RK}^1 &=& 64 \times 10^{-5} MeV \nonumber\\
    T_{LF}^1 &=& 9.44 \pm 0.03 s \nonumber\\
    T_{RK}^1 &=& 21.33 \pm 0.02 s \nonumber\\
    \Delta E_{LF} \times T_{LF} &=& 104 \times 10^{-4} \pm 10^{-4} MeV s \nonumber\\ %0.003
    \Delta E_{RK} \times T_{RK} &=& 137 \times 10^{-4} \pm 2\times 10^{-4}  MeV s \nonumber\\ %0.001
    F_{LF}/F_{RK} &=& \frac{\Delta E_{RK} \times T_{RK} }{\Delta E_{LF} \times T_{LF}^1} = 1.32 \pm 0.01
\end{eqnarray}
Also, observing from the data,
\begin{eqnarray} \label{eq:f_lf_rk_mag_2}
    \Delta E_{LF} (dt=3 \times 10^{-5}) \approx \Delta E_{RK} (dt=5 \times 10^{-5}) &\approx& 32.5 \times 10^{-4} MeV \nonumber\\
    T_{LF}(dt=3 \times 10^{-5})  &=& 3.18 \pm 0.02 s \nonumber\\ %0.0032
    T_{RK}(dt=5 \times 10^{-5})  &=& 4.30 \pm 0.01 s \nonumber\\ %0.0022
    F_{LF}/F_{RK} = \frac{T_{RK}(dt=5 \times 10^{-5})}{T_{LF}(dt=3 \times 10^{-5})} &=& 1.4 \pm 0.1
\end{eqnarray}
Uncertainty of \fromeq{f_lf_rk_mag_2} was taken high due to the approximation. After combining \fromeqs{f_lf_rk_mag_1}{f_lf_rk_mag_2}, $F$ can be calculated as
\begin{equation}
    F_{LF}/F_{RK} = 1.36 \pm 0.05
\end{equation}
Therefore, \textit{Leap-frog} was found to be the better choice as it provided with $1.36 \pm 0.05$ times accuracy in \eB interactions at a given time with respect to \textit{RK4}. However,
\textit{RK4} was promising in situations where decreasing the stepsize, $dt$, is not viable. Therefore both were integrated into \textit{Rhodotron Simulation} for the user to decide.
\textit{RK4-2} renders from these test can be found in \fromfig{mag_rk_render} of \fromapp{example_simulation_runs}.

@@

Considering the $\vecthreeBF{E}$ is parallel to the beam path, potential difference $V$ in the trejectory until \vecNum{-0.753}{0}{0}m is
\begin{eqnarray}
    \Delta V^1 &=& - \int \dotProdBF{E}{ds} \\
             &=& - \int_{-0.753}^{0.753} -2.65616 \times dx \\
             &=& 2.65616 \times 1.506 \\
    \Delta V^1 &=& 4 MV \\
    \Delta E^1 &=& 4 MeV \\
    E_{exitTH}^1 &=& 5 MeV
\end{eqnarray}
In the second test on the other hand, the beam would be injected into a different static uniform electric field, \\
$\vecthreeBF{E}=(0, -5.31232, 0)$ MV/m where $-0.753<x<0.753$ and $-0.753<y<0.753$, $\vecthreeBF{E}=0$ elsewhere.
\begin{eqnarray}
    \Delta V^2   &=& - \int \dotProdBF{E}{ds} \\
                 &=& - \int_{0}^{0.753} -5.31232 \times dy \\
                 &=& 5.31232 \times 0.753 \\
    \Delta V^2   &=& 4 MV \\
    \Delta E^2   &=& 4 MeV \\
    E_{exitTH}^2 &=& 5 MeV
\end{eqnarray}
Therefore, in the both tests, the beam was expected to exit $\vecthreeBF{E}$ with $E_{exitTH} = 5$ MeV. 
To also measure the variance in simulation completion times, $T_{sim}$, set of $10$ runs were completed at the configuration for each $dt$ value.
Results from these tests can be observed in \fromfig{lf_rk2_perp_stat_E_comparison} and \fromfig{lf_rk2_par_stat_E_comparison} of \fromapp{data_graph}.
\begin{figure}[H]
    \centering
    \begin{subfigure}{0.9\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/statE_par.png}
        \caption*{Setup-1, $\textbf{V}_{\parallel}$=4MV}
    \end{subfigure}
    \begin{subfigure}{0.9\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/statE_perp.png}
        \caption*{Setup-2, $\textbf{V}_{\perp}$=4MV}
    \end{subfigure}
    \caption{Render of the test setups. \\ $E_{in}=1$ MeV, $t_{end}=6$ ns}
\end{figure}

@@

    \begin{subfigure}{0.9\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/analiz/90staticE_lf_rk2_dt-Tsim.png}
    \end{subfigure}
    \caption{Comparing Leap-frog, RK4-2 performance on \eE interaction\\ $E_{in}=1$MeV, $\textbf{V}_{\perp}$=4MV, $t_{end}=6$ ns}
    \label{fig:lf_rk2_perp_stat_E_comparison}
\end{figure}
Taking the $dt=10^{-5}$ ns for the reference point as before, the relative performance can be calculated.
\begin{eqnarray*}
    \begin{aligned}
        \Delta E_{LF}^1 &=& 22 \times 10^{-6} MeV \\
        \Delta E_{RK}^1 &=& 16 \times 10^{-6} MeV \\
        T_{LF}^1 &=& 12.40 \pm 0.04 s \\
        T_{RK}^1 &=& 40.08 \pm 0.12 s \\
    \end{aligned}
    \qquad
    \begin{aligned}
        \Delta E_{LF}^2 &=& 15 \times 10^{-5} MeV \\
        \Delta E_{RK}^2 &=& 9.3 \times 10^{-5} MeV \\
        T_{LF}^2 &=& 10.75 \pm 0.04 s \\
        T_{RK}^2 &=& 34.93 \pm 0.39 s 
    \end{aligned}
\end{eqnarray*}
